---
title: "Analyse dienstenprijzen"
author: "Arie Twigt"
date: "`r Sys.Date()`"
output: 
  word_document:
---


# Inleiding

Dit is een analyse van de dienstenprijzen.

# Analyse

## Libraries importeren




```{r}
library(readr)
library(dplyr)
library(zoo)
```




## Importeren bronbestand

Load the csv-file.

Transactions:

```{r}
diensten <- read_delim("data/diensten.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
```


Load the meta-data:

```{r}
metadata_diensten <- read_delim("data/metadata_diensten.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, 
    skip = 11)
```



## Data preparation

-   Deal with the different levels. We have quarterly and yearly.

    -   Remove the yearly rows

-   Some rows contain NA values. These can be removed.

-   The columns containing the prices are characters.

    -   Convert to to numeric data types.



```{r}

numeric_cols <- c("Prijsindex_1", "Kwartaalmutaties_2", "Jaarmutaties_3")

diensten_clean <- diensten %>%
  mutate(ID = as.integer(ID)) %>%
  mutate_at(numeric_cols, funs(as.numeric)) %>%
  filter(!is.na(Kwartaalmutaties_2)) %>%
  mutate(Perioden = as.Date(as.yearqtr(Perioden, format = "%YKW0%q"))) %>%
  mutate(jaar = format(Perioden, "%Y")) %>%
  mutate(kwartaal = quarters(Perioden)) %>%
  mutate(titel = metadata_diensten$Title[match(CPA2008, metadata_diensten$Key)]) %>%
  mutate(titel = gsub("[[:digit:]]+ ", "", titel)) %>%
  filter(titel != "Dienstenprijsindex (DPI)") %>%
  rename_all(.funs = funs(tolower(.))) %>%
  select(-c("cpa2008", "jaarmutaties_3")) %>%
  rename(prijsindex = prijsindex_1) %>%
  rename(kwartaalmutaties = kwartaalmutaties_2) %>%
  select(id, jaar, kwartaal, perioden, titel, prijsindex, kwartaalmutaties)
```
